\textbf{The Illusion Game: A Novel Experimental Paradigm to study Visual Illusions}

### 2022-08-31 manuscript.Rmd
edit

Crucially, this study allowed us to refine the range of task difficulty and illusion strength values in order to maximize the information gain. In most illusions, the task difficulty exhibited a non-linear effect, which is in line with the psychophysics literature on perceptual decisions. One notable result was the illusion effect for the Zöller illusion, which suggested a non-linear pattern. By generating a wider range of illusion strength values, the next study will attempt at clarifying this point.

@@

Using the same recruitment procedure, we recruited 256 participants, from which 6 were identified as outliers and discarded, leading to a final sample of 250 participants (Mean age = 26.5, SD = 7.6, range: [18, 69]; Sex: 48% females, 52% males). See the next study for the full demographic breakdown. We discarded blocks with more than 50% of errors (2.16% of trials) and 0.76% trials based on response time (< 125 ms or > 4 SD above mean).

@@

The best models were $log(diff)*strength$ for Delboeuf;

Because the random-slopes models for RT did not appear to capture well the underlying, likely quadratic, dynamics, we decided not to use add them for the individual scores analysis. Additionally, it would have increased the parameter space from 30 to 90 variables, which was deemed as statistically unreasonable given our sample size).

@@

Future studies should investigate reaction times to find the most appropriate models, and / or use models that integrate errors and reaction time (e.g., drift diffusion models).

@@

>>>>

Crucially, this study allowed us to refine the range of task difficulty and illusion strength values in order to maximize information gain. In most illusions, the task difficulty exhibited a non-linear effect, which is in line with the psychophysics literature on perceptual decisions (*REF*). One notable result was the illusion effect for the Zöller illusion, which suggested a non-linear pattern. By generating a wider range of illusion strength values, the next study will attempt at clarifying this point.

@@

Using the same recruitment procedure, we recruited 256 participants, from which 6 were identified as outliers and discarded, leading to a final sample of 250 participants (Mean age = 26.5, SD = 7.6, range: [18, 69]; Sex: 48% females, 52% males). See the next study for the full demographic breakdown. We discarded blocks with more than 50% of errors (2.16% of trials) and 0.76% trials with extreme response times (< 125 ms or > 4 SD above mean).

@@

The best models were $log(diff)*strength$ for Delboeuf; $sqrt(diff)*strength$ for Ebbinghaus; $log(diff)*log(strength)$ for Rod and Frame; $sqrt(diff)*sqrt(strength)$ for Vertical-Horizontal; .... In all of these models, the effect of illusion strength, task difficulty and their interaction were significant.

For errors, most of the models closely matched their GAMs counterpart, with the exception of Delboeuf for which the GAM suggested a non-monotonic effect of illusion strength with a local minimum at 0.

However, for reaction times, the GAMs suggested a consistent non-linear relationship between the mean RT and illusion strength: as the illusion strength gets very strong, the reaction times are faster. While this is is not surprising (strong illusions are so biasing that it is "easier", i.e., faster, to make the wrong decision), the linear models were not designed to capture this pattern and hence are not good representatives of the underlying dynamics. Because the random-slopes models for RT did not appear to capture well the underlying, likely quadratic, relationship, we decided not to use add them for the individual scores analysis.
<!-- Additionally, it would have increased the parameter space from 20 to 40 variables, which was deemed as statistically unreasonable given our sample size). -->

Though imperfect, we believe that these models capture inter-individual differences more accurately than basic empirical scores (e.g., the total amount of errors, or the average RT). Thus, for each illusion, we have extracted the effect of illusion strength and its interaction with task difficulty (when the illusion incongruent) within each participants. These twenty participant-level scores were subjected to exploratory factor analysis (EFA), which suggested that ...

@@

<!-- This is likely to be caused by the outline. It will be fixed in the next release of the Pyllusion software. -->
Future studies should investigate reaction times to find the most appropriate models, and / or use models that integrate errors and reaction time (e.g., drift diffusion models).

@@

Thus opening the door for new illusions-based paradigms and tasks, we have shown that illusions can be manipulated continuously across several different visual illusions with the parametric illusion generation framework introduced with *Pyllusion* (*REF*). By applying this approach to a reaction-time task, we were able to measure inter-individual scores of objective illusion sensitivity.
